---
title: "Example simulation"
author: "Denis Agniel"
date: December 11, 2016
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Set-up
Setting initial conditionas and generating simulated data:

```{r}
library(fpvc)
TT <- 2*pi
maf <- 0.1
n <- 100
gamma <- 0.75
alpha <- 0.25
beta <- 1
nt <- 10
set.seed(0)

dat <- generateData(n, nt = nt, maf = maf, beta = beta, TT = TT, alpha = alpha, gamma = gamma)
```

## Summarizing longitudinal data

Creating polynomial basis:
```{r}
this.ds <- dplyr::mutate(dat,
                  y1_scaled = c(scale(y1)),
                  y2_scaled = c(scale(y2)))
this.ds <- dplyr::mutate(this.ds,
                         tt2 = scale(tt)^2,
                         tt3 = scale(tt)^3,
                         tt4 = scale(tt)^4,
                         tt5 = scale(tt)^5)
```

Estimating FPCA values:

```{r}
fpc_fit <- fpca(this.ds, ycols = c('y1_scaled', 'y2_scaled'), tcol = 'tt', idcol = 'id', options = NULL)
fpc_scores <- fpc_fit$scores
```

Getting scores for B-spline and polynomial basis functions using
3 basis functions:
```{r}
poly_scores <- polynomialScores(this.ds, df = 3)$ds
bs_scores <- bSplineScores(this.ds, df = 3)$ds
```

Doing the same for re-fitted scores using the eigenfunctions as a basis in a mixed effects model:

```{r}
fpme1 <- fpme(fpc_fit$fpca_fits[[1]], ids = fpc_fit$ids, df = 3)
fpme2 <- fpme(fpc_fit$fpca_fits[[2]], ids = fpc_fit$ids, df = 3)
fpme_scores <- merge(fpme1, fpme2, by = 'id')
```

### Testing using 3 basis functions for each method

```{r}
g_ds <- dplyr::select(this.ds, id, g) 
g_ds <- unique(g_ds)
fpc_ds <- merge(fpc_scores, g_ds, by = 'id')
fpc_test <- fpvc(fpc_ds, xicols = c('xi1.1', 'xi1.2', 'xi1.3',
                                   'xi2.1', 'xi2.2', 'xi2.3'),
                 zcols = 'g', xcols = NULL, model = 'binomial')

refit_ds <- merge(fpme_scores, g_ds, by = 'id')
refit_test <- fpvc(refit_ds, xicols = c('xime.1.x', 'xime.2.x', 'xime.3.x',
                                        'xime.1.y', 'xime.2.y', 'xime.3.y'),
                                        zcols = 'g', xcols = NULL, 
                                        model = 'binomial')

bs_ds <- merge(bs_scores, g_ds, by = 'id')
bs_test <- fpvc(bs_ds, xicols = c('b1.1', 'b1.2', 'b1.3',
                                   'b2.1', 'b2.2', 'b2.3'),
                 zcols = 'g', xcols = NULL, model = 'binomial')

poly_ds <- merge(poly_scores, g_ds, by = 'id')
poly_test <- fpvc(poly_ds, xicols = c('b1.1', 'b1.2', 'b1.3',
                                   'b2.1', 'b2.2', 'b2.3'),
                 zcols = 'g', xcols = NULL, model = 'binomial')
c(fpc_test = fpc_test, 
  refit_test = refit_test,
  bs_test = bs_test,
  poly_test = poly_test)
```
